Analysis2

### 2023-10-20 Lahman Analysis/Rough Analysis 2/Analysis2.qmd
Add simple model (which compiles!) to file
I added a simple model with just `Age` as a predictor so that we could compile something. Before fitting the model, I made a graph for guidance. It would be good to explore this simple model further, including a prior predictive check and a posterior predictive check. At very least, there should be a graph of the model. (Maybe I'll add that next.)

@@

#Add player's age to the "Batting" DF
Batting <- Batting %>% 
  right_join(Lahman::People, by = join_by(playerID)) %>% 
  mutate(Age = yearID - birthYear, .after=yearID) %>%
  #Don't consider the people who don't have a birthday in the database
  filter(!is.na(Age))

@@

>>>>

```{r}
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)
```

@@

#Add player's age to the "Batting" DF
Batting <- Batting %>% 
  right_join(Lahman::People, by = join_by(playerID)) %>% 
  mutate(Age = yearID - birthYear, .after=yearID) %>%
  #Don't consider the people who don't have a birthday in the database
  drop_na(Age)

@@

# Graphical exploration

Before building a model, we explore the data by graphing the home-run proportion with classical binomial standard errors.

```{r}
Batting_summary <- Batting %>% group_by(Age) %>%
  summarize(HR = sum(HR), AB = sum(AB)) %>%
  mutate(prop = HR / AB, se = sqrt(prop*(1-prop)/AB))

ggplot(data = Batting_summary) +
  geom_pointrange(mapping = aes(x = Age, y = prop, ymin = prop - se,
                                ymax = prop + se))
```
Let's graph again, restricting to players aged 20 to 40.

```{r}
ggplot(data = Batting_summary %>% filter(Age >= 20 & Age <= 40)) +
  geom_pointrange(mapping = aes(x = Age, y = prop, ymin = prop - se,
                                ymax = prop + se))
```
Our first model will include a linear term for age and will be restricted to this age range.

# Simple model

```{r}
Batting_summary_restricted <- Batting_summary %>%
  filter(Age >= 20 & Age <= 40) %>% mutate(c.Age = Age - 30)

simple_data <-   list(
  N = nrow(Batting_summary_restricted),
  Age = Batting_summary_restricted %>% pull(c.Age),
  AB = Batting_summary_restricted %>% pull(AB),
  HR = Batting_summary_restricted %>% pull(HR)
)
```

```{r}
glm(cbind(HR, AB - HR) ~ c.Age, data = Batting_summary_restricted, family = binomial)
```

```{r}
fit_simple <- stan(
  file = "HR2_simple.stan",
  data = simple_data
)
```

```{r}
summary(fit_simple, pars = c("beta0", "beta1"))$summary %>% signif(3)
```
```{r}
plot(fit_simple, pars = c("beta0", "beta1"))
```
